# Project Summary

```{r}
#| echo: false
#| output: false

# library(readr) # get data

# library(haven) # Stata data

library(ggplot2) # beautiful graphs

library(dplyr) # data wrangling

library(tidyr) # tidy data

# library(DT) # nice data table

library(tibble) # tibbles

library(countrycode) # manage country data

# library(maps) # maps

library(leaflet) # web maps

# library(sp) # spatial data

library(sf) # simple features

```

```{r}
#| echo: false
#| output: false

global_data <- read_sf("../mapping/shapefiles/wrld_simpl/wrld_simpl.shp") 

# get MICS countries

country <- c("Afghanistan", "Algeria",  "Argentina",  
             "Bangladesh",  "Barbados",  "Belarus",  
             "Belize",  "Benin",  "Bosnia and Herzegovina",  
             "Cameroon",  "Central African Republic",  "Chad",  
             "Democratic Republic of the Congo",  
             "Republic of the Congo",  
             "Costa Rica",  "Cote d'Ivoire",  
             "Dominican Republic",  
             "El Salvador",  "Eswatini",  "Ghana",  
             "Guinea",  "Guinea Bissau",  "Guyana",  
             "Iraq",  "Jamaica",  "Kazakhstan",  
             "Kenya",  "Kosovo",  "Kyrgyzstan",  
             "Laos", "Macedonia",  "Madagascar",  
             "Malawi",  "Mali",  "Mauritania",  
             "Mexico",  "Moldova",  "Mongolia",  
             "Montenegro",  "Nepal",  "Nigeria",  
             "Pakistan",  "Panama",  "Paraguay",  
             "Sao Tome and Principe",  "Senegal",  "Serbia",  
             "Sierra Leone",  "Somalia",  "St. Lucia",  
             "State of Palestine",  "Suriname",  "Thailand",  
             "The Gambia",  "Togo",  "Trinidad and Tobago",  
             "Tunisia",  "Turkmenistan",  "Ukraine",  
             "Uruguay",  "Vietnam",  "Zimbabwe")

# convert to ISO3

country_iso <- countrycode(country, 
                           'country.name', 
                           'iso3c')

# MICS is an sf object that is subset of global_data

MICS <- global_data %>% 
  filter(ISO3 %in% country_iso)
  
```

```{r}
#| fig-cap: "Locations of Countries in MICS"
#| fig-height: 3
#| echo: false
#| eval: true

if (knitr::is_html_output()) {
  
  library(leaflet)
  
  leaflet() %>%
    setView(-25, 25, zoom = 1.25) %>%
    # addTiles() %>%
    addProviderTiles(providers$Esri.WorldImagery) %>%
    addPolygons(data = MICS, 
                weight = 1,
                fillOpacity = .5, 
                color = "#1CABE2", 
                label = MICS$NAME,
                highlightOptions = highlightOptions(color = "red", 
                                                    weight = 2,
                                                    bringToFront = TRUE)) 
}

```

## Introduction

Gender inequality perpetuates harmful norms that justify violence against women and children and is associated with higher rates of family violence. 

Worldwide, parental physical abuse is a common form of family violence that children are exposed to at alarming rates. Parental engagement in physical abuse is linked to negative child outcomes including depression, anxiety, and aggression that may persist into adulthood. Globally, these continuing mental health and aggression problems may have high financial costs, with effects both on social service systems and developing economies. 

Despite the substantial scholarship on parent- and family-level predictors of parent-to-child physical violence, important questions remain about societal-level predictors of parental physical abuse and its associations with young children’s development in developing and transitional countries. 

A further gap in prior literature is the lack of studies that have examined potential moderators such as child age and household economic status in the associations between gender inequality and parental violence against children. 

Using data from over 520,000 families in 57 low- and middle-income countries (LMICs), the current project seeks to address these research gaps by examining the associations of country-level gender inequality and violent social contexts with caregivers’ use of physically abusive behavior and child social-emotional development. We will employ multilevel models using data on parental physical violence against children, family socio-economic characteristics, and children’s social-emotional development from the UNICEF Multiple Indicator Cluster Surveys (MICS) and data on country-level gender inequality and violent social contexts from the United Nations Development Programme on Human Development and the World Health Organization Global Health Observatory. 

## Specific Aims

The specific aims are to: 

1. Examine the associations of gender inequality with parental child physical abuse in LMICs, and the moderating roles of child age and household economic status in these associations. 
2. Examine the associations of violent social norms and crimes with parental physical abuse in LMICs. 
3. Examine the associations of parental physical abuse with child social-emotional development in the context of gender inequality and violent norms and crimes in LMICs, and whether country-level normativeness of physical abuse moderates these associations. 

## Proposed Studies

The proposed studies will advance the understanding of macro-level social and economic indicators that perpetuate caregivers’ physical violence against children in international contexts. Study findings will inform cross-cultural programs and policies that reduce gender disparities and prevent parental physical abuse to promote child social-emotional development across the globe. 

In addition, these studies will provide rigorous research engagement opportunities to undergraduate students and graduate students and strengthen the research environment at the University of Michigan-Flint.