# Simulated Multi-Country Data

This website makes use of simulated data. Data come from 30 hypothetical countries. Data contain measures of a few key aspects of parenting[^parenting] or caregiving that have proven salient in the empirical literature on parenting to date. The outcome is `aggression` against other children.

```{r}
#| label: fig-globe2
#| fig-cap: "Countries of the World"
#| warning: false
#| echo: false
#| fig-height: 3

if (knitr::is_html_output()) {

library(plotly)

plot_geo() %>%
  layout(title = "", 
         geo = list(projection = list(type = 'orthographic'),
                    showland = TRUE,  
                    landcolor = toRGB("#1CABE2"),
                    showcountries = TRUE,
                    showocean = FALSE, 
                    oceancolor = "LightBlue",
                    lataxis = list(showgrid = TRUE,
                                   gridcolor = toRGB("black")),
                    lonaxis = list(showgrid = TRUE,
                                   gridcolor = toRGB("black"))))
}

```

:::{.callout-tip collapse="false"}
## Variable Values

Most variables in this data set are *dichotomous*, and are coded `0/1` with `0 = "no"` and `1 = "yes"`. More descriptive `no/yes` *value labels* could be added in most statistical packages. 
:::

:::{.callout-note collapse="true"}
## Download The Data

* [R](https://github.com/agrogan1/globalfamilies/raw/main/simulate-data/MICSsimulated.RData)
* [Stata](https://github.com/agrogan1/globalfamilies/raw/main/simulate-data/MICSsimulated.dta)
* [SPSS](https://github.com/agrogan1/globalfamilies/raw/main/simulate-data/MICSsimulated.sav)
* [CSV](https://github.com/agrogan1/globalfamilies/raw/main/simulate-data/MICSsimulated.csv)
:::

```{r}
#| code-fold: true
#| code-summary: "Show the code"

load("./simulate-data/MICSsimulated.RData")

```

## Variables and Variable Labels

```{r}
#| code-fold: true
#| code-summary: "Show the code"

labelled::look_for(MICSsimulated)

```

[^parenting]: We use the term parenting throughout this site, but are aware that such parenting may come from biological parents, or from other caregivers.

## A Sample Of The Data

A sample of the data is given below.

```{r}
#| code-fold: true
#| code-summary: "Show the code"

head(MICSsimulated)

```
