# Simulated Multi-Country Data

This website makes use of simulated data. Data come from 30 hypothetical countries. Data contain measures of a few key aspects of parenting[^parenting] or caregiving that have proven salient in the empirical literature on parenting to date. The outcome is `aggression` against other children.

:::{.callout-note collapse="true"}
## Download The Data

* [R format](https://github.com/agrogan1/globalfamilies/raw/main/simulate-data/MICSsimulated.RData)
* [Stata Format](https://github.com/agrogan1/globalfamilies/raw/main/simulate-data/MICSsimulated.dta)
* [SPSS](https://github.com/agrogan1/globalfamilies/raw/main/simulate-data/MICSsimulated.sav)
:::

```{r}
#| code-fold: true
#| code-summary: "Show the code"

load("./simulate-data/MICSsimulated.RData")

```

## Variables and Variable Labels

```{r}
#| code-fold: true
#| code-summary: "Show the code"

labelled::look_for(MICSsimulated)

```

[^parenting]: We use the term parenting throughout this site, but are aware that such parenting may come from biological parents, or from other caregivers.

## A Sample Of The Data

A sample of the data is given below.

```{r}
#| code-fold: true
#| code-summary: "Show the code"

head(MICSsimulated)

```
